---
title: "Correct_colony_depth"
author: "Hugo DENIS"
date: "25/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###############################################################################################
######### Correct colonies depth obtained from dive computer to account for tide level ########
###############################################################################################

#Requires at least
1/ colonies_metadata file with Genotype ID, Collection DateTime, Depth, Reef.name

2/ tide_data file with for each Reef.name,  lowest/highest tides and corresponding times organized as follow
Reef.name , Time1, Time2, Time3, Time4, Level1, Level2, Level3, Level4

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)

#Set to English
Sys.setenv("LANGUAGE"="En")
Sys.setlocale(locale = "C")
```

#Load data
```{r}
#Colony data
path="Data/"
colonies_metadata<-read.csv(file=paste0(path,"RRAP 2022 field data sheets_Feb-Mar_colony.csv"),header=T,check.names = F)

#Tide data
tide_data<-read.csv(paste0(path,"tide_data_ECT1_v3.csv"),header=T,sep=";",dec=".",check.names=F)
names(tide_data)[1]<-"Reef.name"

#Merge the two tables
colonies_metadata<-colonies_metadata %>% merge(tide_data %>% dplyr::select(Reef.name,Time1:Level4),by="Reef.name",all.x=T)

#Convert times to POSIXct
colonies_metadata[,c("Time1","Time2","Time3","Time4")]<-lapply(colonies_metadata[,c("Time1","Time2","Time3","Time4")],function(x)as.POSIXct(paste(colonies_metadata$`Collection date`,x,sep=" "),format="%d/%m/%Y %H:%M:%S"))
```


#### Depth correction of the tide at the time of collection
```{r}
for(i in 1:nrow(colonies_metadata)){
  
  #Avoid missing values
  if(!is.na(colonies_metadata$Depth[i])){
#Take into account the case where time of sampling equals one of tide times   
      if(colonies_metadata[i,"Time"] %in% c(colonies_metadata[i,"Time1"],colonies_metadata[i,"Time2"],colonies_metadata[i,"Time3"],colonies_metadata[i,"Time4"])){
        id_equal=which(c(colonies_metadata[i,"Time1"],colonies_metadata[i,"Time2"],colonies_metadata[i,"Time3"],colonies_metadata[i,"Time4"])==colonies_metadata[i,"Time"])
        colonies_metadata$Tide_adjust[i]=colonies_metadata[i,paste0("Level",as.character(id_equal))] 
      }else{
          #Find Highest/Lowest tide flanking time of collection
      id_inf=max(which(c(colonies_metadata[i,"Time1"],colonies_metadata[i,"Time2"],colonies_metadata[i,"Time3"],colonies_metadata[i,"Time4"])<colonies_metadata[i,"Time"]))
      id_sup=min(which(c(colonies_metadata[i,"Time1"],colonies_metadata[i,"Time2"],colonies_metadata[i,"Time3"],colonies_metadata[i,"Time4"])>colonies_metadata[i,"Time"]))
      time_inf=paste0("Time",as.character(id_inf))
      time_sup=paste0("Time",as.character(id_sup))
      level_inf=paste0("Level",as.character(id_inf))
      level_sup=paste0("Level",as.character(id_sup))
      
      #Find parameters of sinusoid curve passing through those 2 points
      diff=(colonies_metadata[i,level_inf]-colonies_metadata[i,level_sup])
      #Amplitude
      A=abs(diff/2)
      #Period
      b=pi/as.numeric(difftime(colonies_metadata[i,time_sup], colonies_metadata[i,time_inf], units = "min"))
      #Phase
      c=as.numeric(difftime(colonies_metadata[i,if_else(diff>0,time_inf,time_sup)],as.POSIXct(paste0(format(colonies_metadata[i,c("Time1")],format="%Y-%m-%d")," ","00:00:00"),format="%Y-%m-%d %H:%M:%S"),units="min"))-pi/(2*b)
      #Vertical Shift
      d=(colonies_metadata[i,level_inf]+colonies_metadata[i,level_sup])/2
      
      #Find tide height at the time of collection
      timemin=as.numeric(difftime(colonies_metadata[i,"Time"],as.POSIXct(paste0(format(colonies_metadata[i,c("Time1")],format="%Y-%m-%d")," ","00:00:00"),format="%Y-%m-%d %H:%M:%S"),units="min"))
      y=A*sin(b*(timemin-c))+d
  
      colonies_metadata$Tide_adjust[i]=y
      } 
  } else {
    colonies_metadata$Tide_adjust[i]=NA
  }
  
}

#correct depth with the tide 
colonies_metadata$Depth_corrected=round(colonies_metadata$Depth-colonies_metadata$Tide_adjust,1)

#Save as csv with the new column 
write.csv(colonies_metadata %>% dplyr::select(-c(Time1:Level4)),paste0(path,"RRAP 2022 field data sheets_Feb-Mar_colony.csv"),row.names = F)

```

