---
title: "Environment_data_analysis"
author: "Hugo DENIS"
date: "22/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=F}
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(stringr)
library(readr)
library(tidyverse)
library(ggfortify)
library(cowplot)
library(ggpmisc)
library(factoextra)
library(corrplot)

#Set to English
Sys.setenv("LANGUAGE"="En")
Sys.setlocale(locale = "C")
```

#Import environmental data at the colony and site level 

#Note : DHW_max_1year is almost equivalent to DHW at time of collection and DHW_max_6year is almost equivalent to DHW_max so no need to include them

```{r}
path="Data/"
path_out="Figures/"
colony_env=read.csv(paste0(path,"ECT1_Aspat_colonies_environmental_data_24metrics.csv"),header=T)

#Define color palette for sites
mycolors<-c("#1ba3c6" ,"#2cb5c0" ,"#21B087", "#33a65c" ,"#a2b627" ,"#f8b620", "#f89217", "#f06719","#e03426" ,"#f64971", "#fc719e" , "#ce69be", "#7873c0", "#4f7cba")
```

#Plot pearson correlation between those metrics 
```{r}
colony_env=colony_env %>% column_to_rownames(var="Genotype")

cor=cor(colony_env,method="pearson")

png(height=465, width=465, file=paste0(path_out,"SupFigure_8.png"),res=300,units='mm')
corrplot::corrplot(cor, order = 'hclust', addrect = 5,col = COL2('RdYlBu', 10),cl.pos = 'b',tl.cex=2,cl.cex = 2,tl.col='black')
dev.off()
```

#Output tables for supplementary data 
#Save a dataset with "uncorrelated" variables (R<0.7)
```{r}
#Make table for Supplementary Table 9b
#Group variables with pairwise correlation coefficients above 0.7
pw_cor=data.frame()
for(i in 1:nrow(cor)){
  for(j in 1:nrow(cor)){
    if(abs(cor[i,j])>0.7 & i!=j){
      pw_cor=rbind(pw_cor,data.frame(Var1=rownames(cor)[i],Var2=colnames(cor)[j],Cor=cor[i,j]))
    }
  }
}
print(pw_cor)
#write.csv(pw_cor,paste0(path,"Hugo/Data/GBR_env_data/Pairwise_correlations_v5_0.7.csv"),row.names=F)

#List of correlated variables to remove from pairwise correlations
redundant=c("LMM","OM","OMsd","MMM","CF_OM","CF_ss","Chl_a_OM","DHW_freq_sup8","DHW_max","Oxygen_OM","DHW_freq_sup4_6year","Tot_N_OM","PIP_OM")

colony_env_uncor=colony_env %>% dplyr::select(-all_of(redundant))

#%ake table of uncorrelated variables to use in random forest 
colony_env_uncor$Genotype=rownames(colony_env_uncor)
#write.csv(colony_env_uncor,paste0(path,"Hugo/Data/GBR_env_data/ECT1_Aspat_colonies_environmental_data_uncor_v3_0.7.csv"),row.names=F)

#Make supplementary table 9a
colony_env_sub_uncor=colony_env_sub_uncor %>% dplyr::select(-Genotype)
kept_var_table=data.frame(Kept_metric=colnames(colony_env_sub_uncor))

#Find the var with wich the variables that we have excluded were the most correlated
bestcor=data.frame()
for(var in redundant){
  cors=cor[var,]
  bestcor=rbind(bestcor,data.frame(redundant=var,best_cor=names(which.max(abs(cors[!names(cors) %in% c(redundant,var)])))))
}

group_table_sum<-bestcor %>% group_by(best_cor) %>% 
  dplyr::summarize(Variables=paste(redundant,collapse=", "))

kept_var_table=kept_var_table %>% merge(group_table_sum,by.x="Kept_metric",by.y="best_cor",all.x=T)

#write.csv(kept_var_table,paste0(path,"/Hugo/Results/GBR_Env/var_correlated_group_v3_0.7.csv"),row.names=F)
```


#Global PCA on the 11 metrics that are low correlated R<0.9
```{r}
path="Data/"

#Retrieve color scale 
sites_data=read.csv(paste0(path,"GBR_Aspat_sites.csv"),header=T,sep=";")
colnames(sites_data)[1]="Reef.Name"
sites_data=sites_data %>% subset(Reef.Name %in% pheno_data$Site.name)

mycolors<-c("#1ba3c6" ,"#2cb5c0" ,"#21B087", "#33a65c" ,"#a2b627" ,"#f8b620", "#f89217", "#f06719","#e03426" ,"#f64971", "#fc719e" , "#ce69be", "#7873c0", "#4f7cba")
#Order site names by MMM
names(mycolors)<-sites_data$Reef.Name[order(sites_data$MMM)]

colony_env_wide_red=read.csv(paste0(path,"ECT1_Aspat_colonies_environmental_data_uncor_v3_0.7.csv"),header=T)

colony_env_wide_all=read.csv(paste0(path,"ECT1_Aspat_colonies_environmental_data_24metrics.csv"),header=T)

#Add site name info to table 
pheno_data=read.csv(paste0(path,"Phenotypic_trait_per_genotype.csv"),header=T)
colony_env_wide_all=colony_env_wide_all %>% merge(pheno_data %>% distinct(Site.name,Genotype),by="Genotype")

#Arrows name (only R<0.7 + MMM)
arrow_name=c(colnames(colony_env_wide_red),"MMM","Site.name","Genotype")

#Run PCA
env.pca<-prcomp(colony_env_wide_all %>% dplyr::select(-c(Genotype,Site.name)),center=T,scale.=T)

sites=colony_env_wide_all$Site.name

#Main PCA Plot 
pca_plot<-autoplot(env.pca,label=F,label.size=50,colour="transparent",loadings=T,loadings.label=F,loadings.colour="grey",loadings.label.colour = "black",loadings.label.repel = F,loadings.label.size=5) + theme_classic() + geom_point(aes(colour=sites),alpha=0.2,size=2) + theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=20,color="black"),axis.title.y = element_text(size=20,color="black",margin = ggplot2::margin(t = 0, r = 15, b = 0, l = 0)),legend.title = element_text(size=20,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + scale_colour_manual(values = mycolors,name="Site name",labels=names(mycolors)) + guides(colour = guide_legend(override.aes = list(size=5))) 

dat_pca=data.frame(PC1=pca_plot$data$PC1,PC2=pca_plot$data$PC2)
dat_pca$Site.name=sites
sites_pca_barycentre=dat_pca %>% group_by(Site.name) %>% dplyr::summarize(PC1=mean(PC1),PC2=mean(PC2))


#Adjust label position to avoid overlap
sites_pca_barycentre$hjust=c(0.003,0,0.005,-0.002,0,0.004,-0.002,-0.01,-0.002,0.00,0.004,-0.01,-0.006,0)
sites_pca_barycentre$vjust=c(0.002,-0.004,0.006,0,0,0,0,-0.002,-0.003,-0.002,0.001,-0.005,0.004,-0.005)

#Get scaling factor from autoplot
scores <- data.frame(env.pca$x[, 1:2])
scores[] <- lapply(scores, function(x) x / sqrt(sum((x - mean(x))^2)))
loadings <- as.data.frame(env.pca$rotation)[1:2]
scale <- min(max(abs(scores$PC1))/max(abs(loadings$PC1)),
             max(abs(scores$PC2))/max(abs(loadings$PC2))) * 0.8

#Modify loadings names to keep only uncorrelated ones 
PCAloadings <- data.frame(Variables = rownames(env.pca$rotation), env.pca$rotation)
PCAloadings$Variables[!PCAloadings$Variables %in% c(colnames(colony_env_wide_red),"MMM")]=""


pca_plot=pca_plot+ ggrepel::geom_text_repel(size=7.5,aes(x=PC1+hjust,y=PC2+vjust,label=Site.name,colour=Site.name),data=sites_pca_barycentre) + scale_colour_manual(values = mycolors,name="Site name",labels=names(mycolors)) +annotate("text", x = (PCAloadings$PC1)*scale, y = (PCAloadings$PC2)*scale,
     label = PCAloadings$Variables,size=6)

pca_plot$layers[[2]]$aes_params$size <- 0.7
pca_plot$layers[[1]]$aes_params$size <- 50

ggsave(pca_plot,filename=paste0(path_out,"Figure_4.png"),width=14,height=10,dpi=320)
```
