---
title: "Hyperspec_data_analysis"
author: "Hugo DENIS"
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(rstatix)
library(stringr)

library(cowplot)
library(ggpubr)
library(tibble)
library(xlsx)
library(drc)
library(corrplot)
```

#Import data
```{r}
path="Data/"
df_hyper=read.csv(paste0(path,"ECT1_Aspat_Hyperspec_NDVI.csv"),header=T)

#Exclude genotypes that don't have at least 2 fragments per treatment 
Gendelete<-df_hyper %>% group_by(Site.name,Genotype) %>%  dplyr::summarise(Ntreatment=n_distinct(Treatment)) %>% subset(Ntreatment<4)
Gendelete<-c(Gendelete$Genotype,(df_hyper %>% group_by(Site.name,Genotype,Treatment) %>% dplyr::summarise(NFrag=n_distinct(ID)) %>% subset(NFrag<=1))$Genotype)
df_hyper=df_hyper %>% subset(!Genotype %in% Gendelete)
df_hyper=df_hyper %>% arrange(Site.name,Genotype,Treatment)
```

1/ Preliminary tests on NDVI raw measurements

#Control effect of fragment sizes 
```{r}
frag_sizes=read.csv(paste0(path,"Aspat_fragment_sizes.csv"),header=T)
#Some missing fragments from hyperspec are removed from the dataset 
df_hyper2=df_hyper %>% merge(frag_sizes %>% dplyr::select(ID,Npixel),by="ID")

#Fit first model using ML method and use if for backward stepwise selection based on F-tests to compare full and reduced models 
me=lmer(MeanNDVI ~ Npixel + Site.name + Treatment +(1|Tank) + (1|Genotype), data = df_hyper2,REML=F)
summary(me)
#Get confidence intervals
plot(me)
qqnorm(resid(me))
confint(me,level=0.95,method="boot")

```

2/ Computation of ED50 values

#Estimate ED50 for Hyperspec data (Chl and NDVI)
```{r}
metric=c("MeanNDVI")

#Max upper limit is adjusted depending on the metric 
for(m in metric){
  formula=as.formula(paste0(m,"~meanTemp_hold.adj"))
  if(m=="MeanChl"){
    max_upper=0.2
  } else{
    max_upper=0.8
  }
  max_lower=0.05
  for(site in unique(df_hyper$Site.name)){
  model.W2.3<-drm(formula, data=df_hyper %>% subset(Site.name==site), curveid = Genotype, fct=W2.3(names = c('hill', 'max', 'ed50')),upperl=c(NA,max_upper,40),lowerl=c(NA,max_lower,30))
  saveRDS(model.W2.3,file=paste0("Outputs/NDVI/",site,"_",m,"_W2.3_wconstraint.rds"))
}
}

#Same at the site level 
for(m in metric){
  formula=as.formula(paste0(m,"~meanTemp_hold.adj"))
  if(m=="MeanChl"){
    max_upper=0.2
  } else{
    max_upper=0.8
  }
  max_lower=0.05
  for(site in unique(df_hyper$Site.name)){
  site.W2.3<-drm(formula, data=df_hyper %>% subset(Site.name==site), fct=W2.3(names = c('hill', 'max', 'ed50')),upperl=c(NA,max_upper,40),lowerl=c(NA,max_lower,30))
  saveRDS(site.W2.3,file=paste0("Outputs/NDVI/",site,"_",m,"_sitelevel_W2.3_wconstraint.rds"))
}
}

```

#Model selection based on AIC 
```{r}
model_comp=data.frame()
  for(site in unique(df_hyper$Site.name)){
    model_geno=readRDS(paste0("Outputs/NDVI/",site,"_","MeanNDVI","_W2.3_wconstraint.rds"))
    LL.3=drm(formula, data=df_hyper %>% subset(Site.name==site), curveid = Genotype, fct=LL.3(names = c('hill', 'max', 'ed50')),upperl=c(NA,0.8,40),lowerl=c(NA,0.05,30))
    W1.3=drm(formula, data=df_hyper %>% subset(Site.name==site), curveid = Genotype, fct=W1.3(names = c('hill', 'max', 'ed50')),upperl=c(NA,0.8,40),lowerl=c(NA,0.05,30))
    m=data.frame(drc::mselect(model_geno, list(W1.3(), LL.3()), linreg=T, icfct=AIC))
    m$model=row.names(m)
    m$Site.name=site
    m$Metric=me
    row.names(m)=NULL
    m$IC[m$model=="W1.3"]=AIC(W1.3)
    m$IC[m$model=="LL.3"]=AIC(LL.3)
    model_comp=rbind(model_comp,m)
  }

model_comp=model_comp %>% arrange(Site.name,IC)

write.csv(model_comp,paste0("Outputs/NDVI/","NDVI_Modelcomparison_AIC.csv"),row.names=F)
```

#Extract ED values and parameters from models 
```{r}
df_ed=data.frame()
for(m in metric){
  for(site in unique(df_hyper$Site.name)){
    model_W2.3=readRDS(paste0("Outputs/NDVI/",site,"_",m,"_W2.3_wconstraint.rds"))
    
    ed_W2.3=data.frame(ED(model_W2.3,50,interval="delta")) %>% mutate(Site.name=site,Metric=m)
    
    ed_W2.3$Genotype=sapply(strsplit(rownames(ed_W2.3),":"),"[[",2)
    
    #Reformat into single dataframe 
    colnames(ed_W2.3)=c("Estimate","Std.error","Lower","Upper","Site.name","Metric","Genotype")
    ed_W2.3$Model="W2.3"
    
    df_ed=rbind(df_ed,ed_W2.3)
  }
}


df_param=data.frame()
for(site in unique(df_filt_info$Site.name)){
  model_W2.3=readRDS(paste0("Outputs/NDVI/",site,"_",m,"_W2.3_wconstraint.rds"))
  sum_W2.3=broom::tidy(model_W2.3)
  
  #Reformat into single dataframe 
  colnames(sum_W2.3)=c("Term","Genotype","Estimate","Std.error","Statistic","p.value")
  sum_W2.3$Term[sum_W2.3$Term=="ed50"]="e"
  sum_W2.3$Model="W2.3"
  sum_W2.3$Site.name=site
  df_param=rbind(df_param,sum_W2.3)
}

colnames(df_ed)[1]="ED50"
write.csv(df_ed,paste0("Outputs/NDVI/","NDVI_ED50s.csv",row.names=F))

write.csv(df_param,paste0("Outputs/NDVI/","Parameters_W2.3.csv"))

```


3/ Computation of performance retention

#Compute relative ratios from NDVI measurements
```{r}
Hyperspec_ratio=df_hyper %>% group_by(Site.name,Genotype,Treatment) %>% dplyr::summarize(MeanNDVI=mean(MeanNDVI)) %>% dplyr::select(c(Site.name,Genotype,Treatment,MeanNDVI)) %>% tidyr::pivot_wider(names_from = "Treatment",values_from ="MeanNDVI" ) 
colnames(Hyperspec_ratio)[3:6]=c("MMM3","MMM6","MMM9","MMM")
Hyperspec_ratio=Hyperspec_ratio %>% dplyr::mutate(Ratio_NDVI=MMM9/MMM)

Temp_ratio=df_hyper %>% group_by(Site.name,Treatment) %>% dplyr::summarize(mTemp=mean(meanTemp_hold.adj)) %>% tidyr::pivot_wider(names_from = "Treatment",values_from ="mTemp" ) 
colnames(Temp_ratio)[2:5]=c("MMM3","MMM6","MMM9","MMM")
Temp_ratio=Temp_ratio %>% mutate(Ratio_Temp=(MMM9-MMM)/9)
Temp_ratio$MMM_sat=sites_data$MMM
Temp_ratio$MMM_delta=Temp_ratio$MMM-Temp_ratio$MMM_sat

Hyperspec_ratio=Hyperspec_ratio %>% merge(Temp_ratio %>% dplyr::select(Site.name,Ratio_Temp,MMM_delta),by="Site.name") 

Hyperspec_ratio=Hyperspec_ratio %>% mutate(Ratio_NDVI.adj=Ratio_NDVI*Ratio_Temp)

#Account for differences between MMM (treatment ) and MMM site
Hyperspec_site=Hyperspec_ratio %>% group_by(Site.name) %>% dplyr::summarize(mratio=mean(Ratio_NDVI.adj)) 
Hyperspec_site$MMM_delta=Temp_ratio$MMM_delta

anova=aov(mratio~MMM_delta,data=Hyperspec_site)
par(mforw=c(2,2))
plot(anova)
summary(anova)

#Elminate site effect of MMM_delta by using grand mean + MMM_delta residuals 
mod <- lm(Ratio_NDVI.adj~MMM_delta,data=Hyperspec_ratio)
Hyperspec_ratio2 <- broom::augment(mod, data = Hyperspec_ratio) %>%
  mutate(Ratio_NDVI.adj2 = mean(Hyperspec_ratio$Ratio_NDVI.adj) + .resid)
Hyperspec_ratio$Ratio_NDVI.adj=Hyperspec_ratio$Ratio_NDVI.adj2

write.csv(Hyperspec_ratio2,paste0("Outputs/NDVI/","NDVI_ratios_adjusted.csv"),row.names=F)
```


5/ Additional statistics computations to be reported in article

#Average decline in chlorophyll content
```{r}
#Global decline in traits  
sum=df_hyper %>% group_by(Treatment) %>% dplyr::summarise(mNDVI=mean(MeanNDVI
                                                                     ),sd=sd(MeanNDVI)) 

sum=sum%>% mutate(Perc_decline=100*(1-(mNDVI/sum$mNDVI[4])))

a=df_hyper %>% group_by(Genotype,Treatment) %>% dplyr::summarise(mNDVI=mean(MeanNDVI
                                                                     )) %>% pivot_wider(names_from="Treatment",values_from="mNDVI") %>% mutate(Perc_decline=100*(1-(`+9C`/MMM))) %>% subset(Perc_decline>0)

min(a$Perc_decline)
max(a$Perc_decline)

sum(a$Perc_decline<50)/nrow(a)
```




#Save summary of averages per site and treatment
```{r}
sum=df_hyper  %>% group_by(Site.name,Treatment) %>% dplyr::summarise(mNDVI=mean(MeanNDVI)) %>% arrange(Treatment)
#write.xlsx(x=data.frame(sum),file="D:/Hugo/Data/GBR_PAM/NDVI_site_treatment_averages.xlsx",sheetName = "Sheet1")
```

