---
title: "Physiological Analysis"
author: "XXX"
date: "2022-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(cowplot)
library(xlsx)
```

#Import Chla data for subsamples of colonies 
#Lab assays measured on 3 fragments/genotype in MMM-+6C treatments
Chicken : n=30
Pelorus East : n=20
```{r}
path="Data/"
chla=read.csv(paste0(path,"Aspath_chla.csv"),header=T,sep=";",check.names = F)
colnames(chla)[1]="ID"
```

#Regression between NDVI and Chlorophylle a 
```{r}
mplot= chla%>% ggplot(aes(x=1000*Chla_per_dw,y=MeanNDVI)) + geom_point(size=3,alpha=0.6)+ theme_linedraw()+ labs(x="Chlorophyll a (μg/g coral dry weight)",y="NDVI")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7)

log_plot=gphysio%>% ggplot(aes(x=log(1000*Chla_per_dw),y=MeanNDVI)) + geom_point(size=3,alpha=0.6)+ theme_linedraw()+ labs(x="log(Chlorophyll a (μg/g coral dry weight))",y="NDVI")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7)

#Same after aggregation at the genotype level 
mplot2= gphysio%>% group_by(Genotype,Treatment) %>% dplyr::summarise(Chla_per_dw=mean(Chla_per_dw),MeanNDVI=mean(MeanNDVI)) %>% ggplot(aes(x=1000*Chla_per_dw,y=MeanNDVI)) + geom_point(size=3,alpha=0.6)+ theme_linedraw()+ labs(x="Chlorophyll a (μg/g coral dry weight)",y="NDVI")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7)

log_plot2=gphysio %>% group_by(Genotype,Treatment) %>% dplyr::summarise(Chla_per_dw=mean(Chla_per_dw),MeanNDVI=mean(MeanNDVI)) %>% ggplot(aes(x=log(1000*Chla_per_dw),y=MeanNDVI)) + geom_point(size=3,alpha=0.6)+ theme_linedraw()+ labs(x="log(Chlorophyll a (μg/g coral dry weight))",y="NDVI")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7)

both_plot=ggdraw() + draw_plot(mplot,x=0,y=0.5,width=0.5,height=0.5) + draw_plot(log_plot,x=0.5,y=0.5,width=0.5,height=0.5) + draw_plot(mplot2,x=0,y=0,width=0.5,height=0.5) + draw_plot(log_plot2,x=0.5,y=0,width=0.5,height=0.5) + draw_plot_label(label=c("a", "b","c","d"),x=c(0.0, 0.50,0,0.5),c(1,1,0.5,0.5),size = 22)

ggsave(both_plot,filename=paste0("Figures/SupFigure_2.png"),width=12,height=8,dpi=320)
```

#There appears to be a saturation of NDVI values at high chlorophyll content (> 20 ug/g). 
#The following tests aim to assess whether or not this may confound the interpretations. 

1. Test the correlation between the relative decline in NDVI and Chla in the +6 treatment 
```{r}
#Import temperature data
temp_tanks=read.csv(paste0(path,"Aspat_tanks_temperature_adj.csv"),header=T)
temp_tanks$Site_Treatment_Tank=paste(temp_tanks$Site.name,temp_tanks$Treatment,temp_tanks$Tank,sep="_")

chla$Site_Treatment_Tank=paste(chla$Site.name,chla$Treatment,chla$Tank,sep="_")
chla=chla %>% merge(temp_tanks %>% dplyr::select(Site_Treatment_Tank,meanTemp_hold.adj,maxTemp.adj),by="Site_Treatment_Tank")

ratio=chla %>% group_by(Site.name,Genotype,Treatment)%>% dplyr::summarize(MeanNDVI=mean(MeanNDVI),Chla_per_dw=mean(Chla_per_dw),MeanTemp=mean(meanTemp_hold.adj))

df_ratio=data.frame()
for(i in unique(ratio$Genotype)){
  r1=ratio$MeanNDVI[ratio$Genotype==i & ratio$Treatment=="+6C"]/ratio$MeanNDVI[ratio$Genotype==i & ratio$Treatment=="MMM"]
  r2=ratio$Chla_per_dw[ratio$Genotype==i & ratio$Treatment=="+6C"]/ratio$Chla_per_dw[ratio$Genotype==i & ratio$Treatment=="MMM"]
  rtemp=(ratio$MeanTemp[ratio$Genotype==i & ratio$Treatment=="+6C"]-ratio$MeanTemp[ratio$Genotype==i & ratio$Treatment=="MMM"])/6
  df_ratio=rbind(df_ratio,data.frame(Genotype=i,Site.name=unique(ratio$Site.name[ratio$Genotype==i]),MeanNDVI=r1,Chla_per_dw=r2))
}

df_ratio %>% subset(Chla_per_dw<1.2) %>% ggplot(aes(x=Chla_per_dw,y=MeanNDVI)) + geom_point(size=2) + theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5))+ 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7,label.y=0.9)
```

#Compare NDVI measurements to F0 
```{r}
df_filt_info=read.csv(paste0(path,"ECT1_Aspat_FvFm.csv"),header=T,sep=";")
colnames(df_filt_info)[1]="ID"
df_filt_info=df_filt_info %>% subset(qc!=3)

df_hyper=read.csv(paste0(path,"ECT1_Aspat_Hyperspec_NDVI.csv"),header=T,sep=";")

df_hyper=df_hyper %>% merge(df_filt_info %>% dplyr::select(ID,mF),by="ID")

Gendelete<-df_hyper %>% group_by(Site.name,Genotype) %>%  dplyr::summarise(Ntreatment=n_distinct(Treatment)) %>% subset(Ntreatment<4)
Gendelete<-c(Gendelete$Genotype,(df_hyper %>% group_by(Site.name,Genotype,Treatment) %>% dplyr::summarise(NFrag=n_distinct(ID)) %>% subset(NFrag<=1))$Genotype)
df_hyper=df_hyper %>% subset(!Genotype %in% Gendelete)
df_hyper=df_hyper %>% arrange(Site.name,Genotype,Treatment)
df_hyper$Treatment=factor(df_hyper$Treatment,levels=c("MMM","+3C","+6C","+9C"))
#Correlation chlorophyll and F0
df_hyper=df_hyper %>%  group_by(Site.name,Genotype,Treatment)%>% dplyr::summarize(MeanNDVI=mean(MeanNDVI),F0=mean(mF))
df_ratio=data.frame()
for(i in unique(df_hyper$Genotype)){
  r1=df_hyper$MeanNDVI[df_hyper$Genotype==i & df_hyper$Treatment=="+9C"]/df_hyper$MeanNDVI[df_hyper$Genotype==i & df_hyper$Treatment=="MMM"]
  r2=df_hyper$F0[df_hyper$Genotype==i & df_hyper$Treatment=="+9C"]/df_hyper$F0[df_hyper$Genotype==i & df_hyper$Treatment=="MMM"]
  df_ratio=rbind(df_ratio,data.frame(Genotype=i,Site.name=unique(df_hyper$Site.name[df_hyper$Genotype==i]),MeanNDVI=r1,F0=r2))
}

f0_ndvi=df_hyper  %>% ggplot(aes(x=mF,y=MeanNDVI)) + geom_point(size=3,alpha=0.6)+ theme_linedraw()+ labs(x="F0",y="NDVI")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5))+ facet_wrap(.~Treatment) + 
       theme(strip.text.x = element_text(size = 25)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7,label.y=0.9) + ylim(c(0,1)) 

sites_data=read.csv(paste0(path,"GBR_Aspat_sites_data.csv"),sep=";")
colnames(sites_data)[1]="Site.name"

new_palette=c("#1ba3c6" ,"#2cb5c0","#FFF880","#FFE135","#C60012") 

df_ratio$Site.name=factor(df_ratio$Site.name,levels=sites_data$Site.name[order(sites_data$MMM)])

ratio_plot=df_ratio %>% merge(sites_data %>% dplyr::select(Site.name,MMM),by="Site.name") %>% subset(Site.name!="St Crispin")  %>% ggplot(aes(x=F0,y=MeanNDVI)) + geom_point(size=3,alpha=0.6,aes(colour=MMM,group=MMM))+ theme_linedraw()+ labs(x="F0 +9C/MMM",y="NDVI +9C/MMM")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
       theme(strip.text.x = element_text(size = 25)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7,label.y=0.9) + ylim(c(0,1))  +scale_colour_gradientn(colors=new_palette) 

ratio_plot_group=df_ratio %>% merge(sites_data %>% dplyr::select(Site.name,MMM),by="Site.name") %>% subset(Site.name!="St Crispin") %>% group_by(Site.name) %>% dplyr::summarize(F0=mean(F0),MeanNDVI=mean(MeanNDVI),MMM=mean(MMM))  %>% ggplot(aes(x=F0,y=MeanNDVI)) + geom_point(size=5,alpha=0.6,aes(colour=MMM,group=MMM))+ theme_linedraw()+ labs(x="F0 +9C/MMM",y="NDVI +9C/MMM")+ theme(axis.text.x=element_text(size=20,color="black"),axis.text.y=element_text(size=20,color="black"),axis.title.x = element_text(size=18,color="black"),axis.title.y = element_text(size=20,color="black"),legend.title = element_text(size=15,color="black"),legend.text = element_text(size=20,color="black"),legend.position = "none",legend.background = element_rect(fill = "transparent"),
          panel.background = element_rect(colour = "black", size=1.5)) + 
       theme(strip.text.x = element_text(size = 25)) + 
  geom_smooth(method='lm', formula= y~x,size=1.5,colour="black")+
   ggpmisc::stat_poly_eq(formula =y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE,size=7,label.y=0.9) + ylim(c(0,1))  +scale_colour_gradientn(colors=new_palette) 

both_plot=ggdraw() + draw_plot(ratio_plot,x=0,y=0,width=0.5,height=1) + draw_plot(ratio_plot_group,x=0.5,y=0,width=0.5,height=1) + draw_plot_label(label=c("a", "b"),x=c(0.0, 0.50),c(1,1),size = 22)

ggsave(both_plot,filename="C:/Users/Hugo/Documents/Figures/GBR_ED50/F0_NDVI_ratio2.png",width=18,height=10,dpi=320)

```

